\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}

\title{\bf Recognition Science as a Measurement:\\
Axioms, Output, and Test Protocol}
\author{Jonathan Washburn\\
Recognition Physics Institute\\
\texttt{@jonwashburn}\\
Austin, Texas, USA}
\date{\today}

\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{claim}{Claim}

\begin{document}
\maketitle

\begin{abstract}
This note fixes the positioning of Recognition Science (RS): it is not a tunable model but a single deductive instrument that \emph{measures} the unique logical architecture of reality. The instrument is specified by eight axioms; its output is a rigid set of structures and numbers. Mapping that output to SI is done once and never adjusted. The framework stands or falls by whether its single output agrees with all of reality within stated uncertainties.
\end{abstract}

\section*{1. Instrument: the Eight Axioms (Mathematical Core)}
\begin{definition}[Ledger reality]
A recognition event is a double--entry transaction on a universal ledger with immutable positive generator $\delta>0$. Every event posts one debit and one credit; total ledger cost is finite and conserved. Alterations are countable (indivisible ticks).
\end{definition}

\begin{lemma}[Dual balance and symmetry]
Ledger conservation forces a dual pairing of every alteration with its conjugate. Any admissible cost $J$ must obey $J(x)=J(1/x)$.
\end{lemma}

\begin{theorem}[Unique cost functional]
The only analytic cost functional compatible with dual balance, countability, and ledger finiteness is
\[
J(x)=\tfrac12\!\left(x+\frac{1}{x}\right),\qquad x>0.
\]
\end{theorem}

\begin{lemma}[Self-similar update]
The minimal balanced recurrence is $x\mapsto 1+\tfrac{k}{x}$. Countability forces $k\in\mathbb{N}$; global cost minimization selects $k=1$.
\end{lemma}

\begin{theorem}[Golden fixed point]
The unique positive fixed point of $x=1+\tfrac{1}{x}$ is $\varphi=\tfrac{1+\sqrt{5}}{2}$.
\end{theorem}

\begin{theorem}[Stable distinction fixes $d=3$ and $2^3$ ticks]
Stable dual cycles require three spatial dimensions; a spatially complete, ledger--compatible recognition of a minimal voxel uses exactly $2^{3}=8$ atomic ticks.
\end{theorem}

\section*{2. Output: the Dimensionless Measurement}
From the core follow the dimensionless outputs:
\begin{itemize}
  \item universal scale factor $\varphi$,
  \item eight--tick universal clock,
  \item charge--to--rung constructor assigning each field an integer $r$,
  \item path--length $\leftrightarrow$ cost isomorphism, giving masses as exponents of $\varphi$ up to a single universal coherence unit.
\end{itemize}
We write the general mass law as
\[
m \;=\; B\,E_{\mathrm{coh}}\,\varphi^{\,r+f},
\]
with sector factor $B$ fixed by channel multiplicity, integer rung $r$ fixed by gauge charges via the minimal path, and residue $f$ a determined, parameter--free renormalization integral. No continuous parameters appear.

\section*{3. Mapping to SI (once)}
The SI bridge is fixed once by identifying:
\[
c \;\; {\rm finite\ and\ maximal},\qquad
\lambda_{\rm rec}=\sqrt{\frac{\hbar G}{c^3}},\qquad
E_{\mathrm{coh}} \;{\rm sets\ the\ energy\ unit}.
\]
After this identification there are no further dials. All predictions are locked.

\section*{4. Test Protocol (Falsifiable Now)}
\begin{itemize}
  \item \textbf{Large--scale structure:} RS predicts a scale--dependent linear growth factor $D(a,k)$ that suppresses shear at arcminute scales by a fixed percent. Failure to see the suppression within stated uncertainty falsifies the instrument.
  \item \textbf{Inflation imprints:} A small, fixed tensor ratio and a scalar tilt near $1-2/N$ follow from the RS potential family. A confirmed deviation beyond the RS window falsifies the instrument.
  \item \textbf{Mass ladder:} Given only charges, the constructor and a single $E_{\mathrm{coh}}$ reproduce the pole masses. A systematic failure beyond error propagation falsifies the instrument.
\end{itemize}

\section*{5. What Counts as Success}
RS succeeds if, with the \emph{single} SI mapping above, the locked outputs simultaneously match all relevant observations within propagated uncertainties. RS fails if any headline prediction disagrees at a statistically decisive level. There is no retuning step.

\end{document}


